---
title: "RLoopFusion Examples"
author: "Nick Ulle"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RLoopFusion}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette presents a simple example illustrating how RLoopFusion is used,
and a more sophisticated example that demonstrates the power of the package.

```{r, results = "hide"}
library(RLoopFusion)
```

## Example 1

Consider the following code.

```{r, results = "hide"}
ex1 = quote({
  a = b = c = d = 0
  for (i in 2:10)
    a = i
  for (i in 1:10)
    b = i
  for (i in 1:10)
    c = i
  for (i in 1:10)
    d = d + 1
})
```

The `quote` function tells R to parse but not evaluate the code.
This is done so that loop fusion can be applied.
In this code, only the two loops in the middle should be fused.
The first loop has a different header from the others,
and the last loop is serial whereas all the other loops are parallel.
We can apply loop fusion with the `loop_fusion` function.

```{r}
loop_fusion(ex1)
```

This gives the correct result.

The fusion graph can be viewed with the `fusion_graph` function.

```{r, results = "hide"}
graph = fusion_graph(ex1)
```

In order to plot the graph, first load the `graph` and `Rgraphviz` libraries.
The graph can then be plotted by calling the `plot` function on the `graph`
element of the output.

```{r}
library(graph)
library(Rgraphviz)
plot(graph$graph)
```

Parallel, serial, and block nodes are denoted by "p", "s", and "n", respectively.
The plot does not indicate which edges are fusion-preventing.
In this example, only the edges out of "p2" are fusion-preventing.
This can be viewed in the edge data for the graph.

```{r}
edge_attr(graph$graph, "prevent_fusion")
```

Finally, dependence information can be computed with the `collect_deps` function.
```{r}
collect_deps(ex1)
```
This function can compute dependence information for any code object,
not just code blocks.

## Example 2

Now we consider a more challenging example. The code is shown below.

```{r}
ex2 = quote({
  n = 42

  x = 0
  for (i in 1:n) {
    # A parallel loop.
    x = i
  }

  a = b = 1
  for (i in 1:n) {
    # A serial loop that computes Fibonacci numbers.
    fib = a + b
    a = b
    b = fib
  }

  y = 0
  for (i in 1:n) {
    # A parallel loop that depends on the Fibonacci loop.
    y = fib
  }
})
```

The first and last loop are parallel, while the middle loop is serial.
Fusion is complicated by the fact that the last loop depends on the middle loop.
The loop fusion algorithm correctly fuses the first and last loop, then places
the new loop _after_ the serial loop.

```{r}
loop_fusion(ex2)
```

This example emphasizes the power of Kennedy and McKinley's algorithm
(in fact, it approximates the example they present as Figure 1).
If parallel loop fusion was not done separately from serial loop fusion,
the first two loops might end up in the same partition, and the algorithm
would fail to fuse the two parallel loops. The fusion graph is shown below.

```{r, echo = FALSE}
plot(fusion_graph(ex2)$graph)
```